Making use of spatially biased variables in ecosystem condition accounting – a GIS based workflow

### 2024-06-06 manuscript/Manuscript.qmd
fix Moen ref

We used the entire national data set for assessing the relationship between the HIA levels and the indicator values.
We then subsetted the national data set and chose three municipalities in south Norway to test out the indicators.
The municipalities differ in several aspects, such as the amount of mire area, the total area surveyed, and the prevalence of infrastructure (@tbl-munis, @fig-studyLocation2, S3, S4).
The nature type polygons with the indicator values were intersected with the HIA map and a map of municipal outlines in Norway.
The relationship between the HIA classes and indicator values was examined visually.
For each HIA and municipality combination we then created a probability distribution for the area weighted mean of the indicator values using Bayesian updating of a uniform prior between 0 and 1, informed by the standard deviation (SD) of the indicator values in the national data set (@fig-workflowExample C; Appendix C, line 138 <!--# update line number manually when re-rendering appendix C -->).
We refer to this step as the first of two *aggregation* steps (Figure S1).
The resulting distributions are assumed to be normally distributed, and we therefore simply carried the mean and SD from the posterior distributions over to individual polygons in the ecosystem delineation map, for each HIA class separately.
This step is referred to as *spreading* the data (Figure S1).
For each polygon we then sampled random numbers from a normal distribution with this same mean and SD.
The number of m^2^ dictated the number of samples for each polygon, thus ensuring that large polygons ended up counting more towards the indicator value in the given municipality.
We then drew 1000 random values from this vector of possible indicator values and created a probability distribution for the indicators for each municipality (the second aggregation step; Figure S1).
When there were no indicator values for a given HIA class, we ignored that class also in the municipal estimate.

>>>>

We used the entire national data set for assessing the relationship between the HIA levels and the indicator values.
We then subsetted the national data set and chose three municipalities in south Norway to test out the indicators.
The municipalities differ in several aspects, such as the amount of mire area, the total area surveyed, and the prevalence of infrastructure (@tbl-munis, @fig-studyLocation2, S3, S4).
The nature type polygons with the indicator values were intersected with the HIA map and a map of municipal outlines in Norway.
The relationship between the HIA classes and indicator values was examined visually.
For each HIA and municipality combination we then created a probability distribution for the area weighted mean of the indicator values using Bayesian updating of a uniform prior between 0 and 1, informed by the standard deviation (SD) of the indicator values in the national data set (@fig-workflowExample C; Appendix C, line 134 <!--# update line number manually when re-rendering appendix C -->).
We refer to this step as the first of two *aggregation* steps (Figure S1).
The resulting distributions are assumed to be normally distributed, and we therefore simply carried the mean and SD from the posterior distributions over to individual polygons in the ecosystem delineation map, for each HIA class separately.
This step is referred to as *spreading* the data (Figure S1).
For each polygon we then sampled random numbers from a normal distribution with this same mean and SD.
The number of m^2^ dictated the number of samples for each polygon, thus ensuring that large polygons ended up counting more towards the indicator value in the given municipality.
We then drew 1000 random values from this vector of possible indicator values and created a probability distribution for the indicators for each municipality (the second aggregation step; Figure S1).
When there were no indicator values for a given HIA class, we ignored that class also in the municipal estimate.
